 Casting multiple shadows: interactive data visualisation with tours and embeddings

### 2020-02-06 paper.Rmd
more words in intro

# Related Work

## Embedding methods

We discuss the use of

- PCA
- tSNE
- UMAP
- PHATE

>>>>

As part of a visualization workflow, it's important to consider the perception 
and interpretation of embedding methods as well. @Sedlmair2013-pn showed that 2D
scatter plots were mostly suffucient for detecting class seperation, however
they noted that often multiple embeddings were required. For the task of cluster
identification, @Lewis2012-ai showed experimentally that novice
users of non-linear embedding techniques were more likely to consider clusters
of points on a 2d scatter plot to be the result of a spurious embedding
compared to advanced users who were aware of the inner workings of the embedding
algorithm. @Wattenberg2016-ji noted that in the case of t-distributed stochastic
neighbourhood embeddings (t-SNE) it was important for users to modify algorithm
parameters in order to preserve global structure in the low dimensional
embedding.

Embedding methods can be classified into two broad groups: linear
and non-linear methods. Linear methods perform a linear transformation
of the data; one example is principal components analysis (PCA)
which performs an eigendecomposition of the estimated sample covariance matrix.
The eigenvalues are sorted in decreasing order and represent the variance
explained by each component (eigenvector). A common approach to deciding on
the number of principal components to retain is to plot the proportion of variance explained by each component and choose a cut-off. Non-linear methods
generally perform preprocessing on the high-dimensional data such as generating
a neighborhood graph and perform transformations on the preprocessed form. We
restrict our attention to three methods that are commonly used in
high-throughput biology (though this is by no means exhaustive): t-SNE, 
uniform manifold alignment and projection (UMAP), and PHATE.

Clearly, finding an appropriate embedding is a difficult and somewhat
poorly defined problem. We propose a more pragmatic workflow inspired by (wickham blindfold paper) by incorporating
interactive graphics and tours with embeddings that allows users to see a
global overview of their high dimensional data and assists them with cluster
orientation tasks. This workflow is incorporated into an R package
called `liminal` (Available: https://github.com/sa-lee/liminal).
